---
title: "Performance Summary"
output:
  html_document: default
vignette: >
  %\VignetteIndexEntry{Performance Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

This vignette provides a performance summary of running workflow_cellxgene_h5ad.Rmd and workflow_cellxgene_h5ad_DelayedArray_parameters.Rmd. 

## Time and resources spent in the workflow

CPU usage, memory usage and running time are compared between DelayedArray vs dgcMatrix and medium vs large size file. 240 GB memory and 3 cpus were used. Medium: number of cells: 94256, large: number of cells: 168860

After running workflow Rcollectlworkflowh5ad in Terra workspace, submission ids could be used as input to generate three csv files CPU_mean.csv, MEM_mean.csv and Time.csv with the R code below. 

```{r, message=FALSE, warning=FALSE}
## generate comparison csv files
library(AnVIL)
library(Rcollectl)
library(Rcollectlworkflowh5ad)
avworkspace("bioconductor-rpci-yubo/Rcollectlworkflowh5ad")

inputTribble <- tribble(
  ~name, ~submissionId,
  "dgCMatrix_medium", "17efa80e-3ef1-429d-8359-58c7cd653310",
  "DA_medium", "2f0cc8ee-ae53-44e6-a6c0-31e7089f0d42",
  "dgCMatrix_large", "c3c147eb-2ff8-4cf0-ba55-5708fa9c0ae5",
  "DA_large", "8be18771-f06d-4256-ac13-6566a3053f14"
)
summary_result_dataset(inputTribble)
```

Three csv files are generated in default R work directory, and need to be moved to folder csv in vignette directory.

CPU mean
```{r, echo = FALSE, message=FALSE}
read.csv("csv/CPU_mean.csv")
```

Memory max
```{r, echo = FALSE, message=FALSE}
read.csv("csv/MEM_mean.csv")
```

Time
```{r, echo = FALSE, message=FALSE}
read.csv("csv/Time.csv")
```

## Workflow running performance of DelayedArray with different parameters setup

When reading a H5AD file as a SingleCellExperiment object with assays loaded as DelayedArray, the set up of auto block size and number of cpus impact workflow running performance. Running time, memory usage and cpu usage are evaluated here, and workflow step variance modeling (function scran::modelGeneVarByPoisson) is taken as example.

```{r, echo = TRUE, message=FALSE, eval=FALSE}
# autoBlockSize
library(Rcollectl)
times_delayed_block_size <- lapply(67108864L * 2L^(0:5), function(i) {
    gc()
    setAutoBlockSize(i)
    bpp <- SerialParam()
    system.time({
    	id <- cl_start(paste0("AutoBlockSize_", i))
        dec_delay <- modelGeneVarByPoisson(
            sce_delay, block=sce_delay$sample, BPPARAM=bpp
        )
        cl_stop(id)
    })
})

# cpu
setAutoBlockSize(.Machine$integer.max)
times_delayed <- lapply(1:5 * 2L, function(i) {
    message(i)
    bpp <- MulticoreParam(i)
    system.time({
        id <- cl_start(paste0("cpu_", i))
        dec_delay <- modelGeneVarByPoisson(
            sce_delay, block=sce_delay$sample, BPPARAM=bpp
        )
        cl_stop(id)
    })
})
```

After running workflow workflow_h5ad_DelayedArray_parameters in Terra workspace, submission ids could be used as input to generate six csv files ended with autoBlockSize.csv and six ended with cpu.csv, these could be combined to one file for different autoBlockSize and one for different cpus using the R code below. 

```{r, message=FALSE, warning=FALSE}
## generate csv files for different autoBlockSize and cpu
library(AnVIL)
library(Rcollectl)
library(Rcollectlworkflowh5ad)
avworkspace("bioconductor-rpci-yubo/Rcollectlworkflowh5ad")

submissionId <- "8ca42a70-d2e9-421f-9f51-f42601348a82"
summary_result(submissionId)

submissionId <- "35280de1-42d8-492b-aa8c-5feff984bffa"
summary_result(submissionId)
```

Two csv files are generated in default R work directory for every submission id, one ended with _autoBlockSize.csv, and the other _cpu.csv, then both need to be moved to folder csv in vignette directory.

File url: https://datasets.cellxgene.cziscience.com/0c6f2c74-8d27-4c1f-b63a-a766bc9f9490.h5ad, cell count: 19361

autoBlockSize
```{r, echo = FALSE, message=FALSE}
read.csv("csv/8ca42a70-d2e9-421f-9f51-f42601348a82_autoBlockSize.csv")
```

Number of cpus
```{r, echo = FALSE, message=FALSE}
read.csv("csv/8ca42a70-d2e9-421f-9f51-f42601348a82_cpu.csv")
```

File url: https://datasets.cellxgene.cziscience.com/e0102be9-1756-46a9-aed1-4b2f03aade3c.h5ad, cell count: 94256

autoBlockSize
```{r, echo = FALSE, message=FALSE}
read.csv("csv/35280de1-42d8-492b-aa8c-5feff984bffa_autoBlockSize.csv")
```

Number of cpus
```{r, echo = FALSE, message=FALSE}
read.csv("csv/35280de1-42d8-492b-aa8c-5feff984bffa_cpu.csv")
```

## `R` session information

```{r session_info, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```
