---
title: "Workflow running performance of DelayedArray with different parameters setup"
output:
  html_document: default
params:
  knitr_eval: FALSE
  file_url: "https://datasets.cellxgene.cziscience.com/0c6f2c74-8d27-4c1f-b63a-a766bc9f9490.h5ad"
  sample: "Name"
  mem_gb: 120
vignette: >
  %\VignetteIndexEntry{Workflow running performance of DelayedArray with different parameters setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, eval = params$knitr_eval)
```

This vignette explores when reading a H5AD file as a SingleCellExperiment object with assays loaded as DelayedArray, the impact of parameters (auto block size and number of cpus) set up to workflow running performance. Running time, memory usage and cpu usage are evaluated here, and workflow step variance modeling (function scran::modelGeneVarByPoisson) is taken as example.

## Set up parameters

```{r, echo = TRUE, message=FALSE, eval=TRUE}
mem_gb <- params$mem_gb
```

```{r, echo = TRUE, message=FALSE, eval=TRUE}
file_url <- params$file_url
```

Column to analyze needs to be specified from colData of the dataset.

```{r, echo = TRUE, message=FALSE, eval=TRUE}
sample <- params$sample
```

```{r, echo = TRUE, message=FALSE, eval=TRUE}
## file url
file_url

## column selected for analysis
sample

## memory 
sprintf("%d GB", mem_gb) 
```

## Download h5ad file

```{r file_download, echo = TRUE, message=FALSE}
library(Rcollectl)
library(cellxgenedp)
library(dplyr)

## download h5ad file
h5ad_file <- files(db()) |>
  filter(url == file_url) |> files_download(dry.run = FALSE)
```

## Step data preparation 

```{r data_preparation, echo = TRUE, message=FALSE, warning=FALSE}
## import h5ad file
library(zellkonverter)
library(HDF5Array)
sce <- readH5AD(h5ad_file, use_hdf5 = TRUE, reader = "R")

## pre-process data
library(SingleCellExperiment)
sce$sample <- colData(sce)[, sample]

library(EnsDb.Hsapiens.v86)
rowData(sce)$Chr <- mapIds(EnsDb.Hsapiens.v86, keys=rownames(sce), column="SEQNAME", keytype="GENEID")

names(assays(sce)) <- "counts"
```

## set auto block size

```{r setAutoBlockSize, echo = TRUE}
library(DelayedArray)
setAutoBlockSize(size=min(.Machine$integer.max, mem_gb*1e+9))
```

## Step quality control

```{r quality_control, message=FALSE}
library(BiocParallel)
library(scater)

bpp <- MulticoreParam(as.integer(3))

sce <- unfiltered <- addPerCellQC(sce, BPPARAM=bpp, subsets=list(Mito=which(rowData(sce)$Chr=="MT")))
qc <- quickPerCellQC(colData(sce), batch=sce$sample, sub.fields="subsets_Mito_percent")
sce <- sce[,!qc$discard]

## Normalization
sce <- logNormCounts(sce, size_factors = sce$sum)
sce_delay <- sce
```

## AutoBlockSize

With different autoBlockSize set up and given memory and cpu, Rcollectl is used to record computing resources usage.

```{r AutoBlockSize, message=FALSE}
library(scran)

times_delayed_block_size <- lapply(67108864L * 2L^(0:5), function(i) {
    gc()
    setAutoBlockSize(i)
    bpp <- SerialParam()
    system.time({
    	  id <- cl_start(paste0("AutoBlockSize_", i))
        dec_delay <- modelGeneVarByPoisson(
            sce_delay, block=sce_delay$sample, BPPARAM=bpp
        )
        cl_stop(id)
    })
})
times_delayed_block_size
```

## CPU

With different number of cpus set up and given memory and autoBlockSize, Rcollectl is used to record computing resources usage.

```{r CPU, message=FALSE}
library(scran)
setAutoBlockSize(.Machine$integer.max)

times_delayed_cpu <- lapply(1:6 * 2L, function(i) {
    message(i)
    bpp <- MulticoreParam(i)
    system.time({
        id <- cl_start(paste0("cpu_", i))
        dec_delay <- modelGeneVarByPoisson(
            sce_delay, block=sce_delay$sample, BPPARAM=bpp
        )
        cl_stop(id)
    })
})
times_delayed_cpu
```

## `R` session information

```{r session_info, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```
